---
title: "Weather Project"
author: "Austin Shih"
date: "2022-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  - Weather data base was manually created

 - Over 2021, does the time of sunrise have an effect on the average temperature, humidity, and wind speed per month in Nashville, TN? 
 - Is the effect the same for Cape Town, South Africa, which is in the Southern Hemisphere?

```{r}
library(tidyverse)
library(scales)
library(ggridges)
wthr <- read_csv('../data/WeatherData.csv')

wthr <- wthr %>%
  mutate(month = factor(Date, levels = month.abb))

wthr %>%
  ggplot(aes(x = month, y = AvgTemp, color = Location)) + 
  geom_point() + 
  theme_ridges()

wthr %>%
  ggplot(aes(x = month, y = AvgHum, color = Location)) + 
  geom_point() + 
  theme_ridges()

wthr %>%
  ggplot(aes(x = month, y = AvgWindSpd, color = Location)) + 
  geom_point() + 
  theme_ridges()

wthr %>%
  ggplot(aes(x = AvgSunrise, y = AvgTemp, color = Location)) + 
  geom_point() + 
  geom_smooth(method = lm, se = F) + 
  labs(title = 'Average Time of Sunrise vs. Average Temperature', 
       x = 'Average Time of Sunrise', 
       y = 'Average Temperature (F)') + 
  theme_ridges()

wthr %>%
  ggplot(aes(x = AvgSunrise, y = AvgHum, color = Location)) + 
  geom_point() + 
  geom_smooth(method = lm, se = F) + 
  labs(title = 'Average Time of Sunrise vs. Average Humidity', 
       x = 'Average Time of Sunrise', 
       y = 'Average Humidity (%)') + 
  theme_ridges()

wthr %>%
  ggplot(aes(x = AvgSunrise, y = AvgWindSpd, color = Location)) + 
  geom_point() + 
  geom_smooth(method = lm, se = F) + 
  labs(title = 'Average Time of Sunrise vs. Average Wind Speed', 
       x = 'Average Time of Sunrise', 
       y = 'Average Wind Speed (mph)') + 
  theme_ridges() 
```

> My experiment showed that there is a strong correlation between time of sunrise and temperature. 
> As the time of sunrise becomes later, the average temperature gets lower. 
> However, there doesn’t seem to be any correlation between time of sunrise and humidity and wind speed. 
> Most fluctuations and difference in measurements between the two cities are mostly due to geographical differences rather than difference in hemispheres. 
